Path Analysis Data Development - Chapter 2

### 2016-06-27 sem-data-development.Rmd
Add Julian day and week of year to dataframe for path model.

### Save data with BUSH01 excluded
```{r save data}
stems <- as.tbl(stems)
stems
# stems <- select(stems, -X) 
# unique(filter(stems, plot_status == "Abandoned")$plotid)
stems <- stems %>% filter(plotid != "bush01", plotid != "ponti01", 
                                plotid != "sweet01")
stems$plotid <- droplevels(stems$plotid)
# write.csv(plots_env, "analysis/data/env_gis_variables.csv", row.names = F)
# write.csv(rainy_season_ppt, "analysis/data/rain_data_long.csv", row.names = F)
# write.csv(rs_temps, "analysis/data/rs_temps_data.csv", row.names = F)
# write.csv(summer_temps, "analysis/data/summer_temps_data.csv", row.names = F)
# write.csv(stems, "analysis/data/stems_noBUSH01.csv", row.names = F)
# rm(plots_temps); rm(plots_rain)
```

@@

## Examine the bay laurel data at individual and plot levels
```{r examine umca data}
bay_laurel <- filter(stems, species == "umca") %>% 
      select(plotid, cluster:stem_status, umca_slc, notes:dbh_entry)
bay_laurel$species <- droplevels(bay_laurel$species)
summary(bay_laurel) # This has stem-level data on bay laurel.
length(unique(bay_laurel$plotid))# 183 plots with bay laurel
# bay_laurel <- filter(bay_laurel, plot != "BUSH01")

@@

I can use the stem-level data to calculate plot level leaf count variables (sum, average) for each plot and year.
```{r calculate plot level leaf counts}
live_umca <- bay_laurel %>% filter(stem_status == "Alive")
filter(live_umca, is.na(umca_slc))
# Spot checking some of these shows that NAs are due to missing data, i.e. they should not be changed to zeroes
umca_plots <- left_join(
      umca_plots, 
      live_umca %>% 
            group_by(plotid, year) %>% 
            summarise(avg_lfct = mean(umca_slc, na.rm = T),
                      tot_lfct = sum(umca_slc, na.rm = T)),
      by = c("plotid", "year"))
summary(umca_plots)
```

@@

qusp_stems <- stems %>% 
      select(plotid, year, tag, species, canker, stem_status, infection_confirmed_1plus, infection_confirmed_2plus) %>% 
      filter(species == "quag" | species == "quke" | species == "quch" |
                   species == "quwi", stem_status == "Alive")

@@

# Add bay laurel infection data to oak plots
names(umca_plots)
oak_sod <- left_join(oak_sod, 
                     select(umca_plots, plotid, sample_year, uninfected_bay_ct, infected_bay_ct, tot_bay, tot_lfct), 
                     by = c("plotid", "sample_year"))
summary(oak_sod) # the NA's for the bay laurel variables are legitimately zeroes b/c there are no bay in some of these plots
filter(oak_sod, is.na(tot_bay), sample_year != 2003)
# Change NAs of bay laurel in oak only plots to zeroes
oak_sod$uninfected_bay_ct[is.na(oak_sod$uninfected_bay_ct)] <- 0
oak_sod$infected_bay_ct[is.na(oak_sod$infected_bay_ct)] <- 0
oak_sod$tot_bay[is.na(oak_sod$tot_bay)] <- 0

@@

# summary(filter(oak_sod, is.na(rain_tot_psm))) # all obs after 2010
# summary(filter(oak_sod, is.na(us.J.2011)))
unique(filter(oak_sod, is.na(us.J.2011))$plotid)# ann06, mroth03, sumtv02
# ann06 & mroth03 are legitimate zeroes for understory richness in 2011
# i must have excluded mcneil03 from the diversity calculations
# sumtv02 was abandoned after 2008, so it should be an NA for later years

>>>>

### Save data with BUSH01 excluded
```{r save data}
stems <- as.tbl(stems)
stems
# stems <- select(stems, -X) 
# unique(filter(stems, plot_status == "Abandoned")$plotid)
stems <- stems %>% filter(plotid != "bush01", plotid != "ponti01", 
                                plotid != "sweet01")
stems$plotid <- droplevels(stems$plotid)
stems$date <- as.Date(stems$date)
library(lubridate)
stems$julian_day <- yday(stems$date)
stems$month <- month(stems$date)
stems$week <- week(stems$date)
summary(stems)
detach("package:lubridate", unload=TRUE)
# write.csv(plots_env, "analysis/data/env_gis_variables.csv", row.names = F)
# write.csv(rainy_season_ppt, "analysis/data/rain_data_long.csv", row.names = F)
# write.csv(rs_temps, "analysis/data/rs_temps_data.csv", row.names = F)
# write.csv(summer_temps, "analysis/data/summer_temps_data.csv", row.names = F)
# write.csv(stems, "analysis/data/stems_noBUSH01.csv", row.names = F)
# rm(plots_temps); rm(plots_rain)
```

@@

```{r plot visit dates}
plot_visit_dates <- stems %>% 
      select(plotid, year, julian_day:week) %>% 
      group_by(plotid, year) %>% 
      summarise(julian_day = mean(julian_day), month = mean(month), week = mean(week))
plot_visit_dates <- ungroup(plot_visit_dates)
```

## Examine the bay laurel data at individual and plot levels
```{r examine umca data}
bay_laurel <- filter(stems, species == "umca") %>% 
      select(plotid, cluster:stem_status, umca_slc, notes:week)
bay_laurel$species <- droplevels(bay_laurel$species)
summary(bay_laurel) # This has stem-level data on bay laurel.
length(unique(bay_laurel$plotid))# 183 plots with bay laurel
# bay_laurel <- filter(bay_laurel, plot != "BUSH01")

@@

I can use the stem-level data to calculate plot level leaf count variables (sum, average) for each plot and year.
```{r calculate plot level leaf counts}
live_umca <- bay_laurel %>% filter(stem_status == "Alive")
filter(live_umca, is.na(umca_slc))
# Spot checking some of these shows that NAs are due to missing data, i.e. they should not be changed to zeroes
umca_plots <- left_join(
      umca_plots, 
      live_umca %>% 
            group_by(plotid, year) %>% 
            summarise(avg_lfct = mean(umca_slc, na.rm = T),
                      tot_lfct = sum(umca_slc, na.rm = T)),
      by = c("plotid","year"))
umca_plots <- left_join(
      umca_plots,
      plot_visit_dates,
      by = c("plotid","year")
)
summary(umca_plots)
```

@@

qusp_stems <- stems %>% 
      select(plotid, year, tag, species, canker, stem_status, infection_confirmed_1plus, infection_confirmed_2plus, dbh, dbh_entry, julian_day:week) %>% 
      filter(species == "quag" | species == "quke" | species == "quch" |
                   species == "quwi", stem_status == "Alive")

@@

# Add bay laurel infection data to oak plots
names(umca_plots)
oak_sod <- left_join(oak_sod, 
                     select(umca_plots, plotid, sample_year, uninfected_bay_ct:tot_lfct ), 
                     by = c("plotid", "sample_year"))
summary(oak_sod) # the NA's for the bay laurel variables are legitimately zeroes b/c there are no bay in some of these plots
oak_sod <- left_join(
      oak_sod,
      plot_visit_dates,
      by = c("plotid","sample_year"="year"))

filter(oak_sod, is.na(tot_bay), sample_year != 2003)
# Change NAs of bay laurel in oak only plots to zeroes
oak_sod$uninfected_bay_ct[is.na(oak_sod$uninfected_bay_ct)] <- 0
oak_sod$infected_bay_ct[is.na(oak_sod$infected_bay_ct)] <- 0
oak_sod$tot_bay[is.na(oak_sod$tot_bay)] <- 0

@@

# summary(filter(oak_sod, is.na(rain_tot_psm))) # all obs after 2010
# summary(filter(oak_sod, is.na(us.J.2011)))
unique(filter(oak_sod, is.na(us.J.2011))$plotid)# ann06, mroth03, sumtv02
# ann06 & mroth03 are legitimate zeroes for understory richness in 2011
# i must have excluded mcneil03 from the diversity calculations
# sumtv02 was abandoned after 2008, so it should be an NA for later years
filter(oak_sod, is.na(hrs_abv25rs))
oak_sod$hrs_abv25rs[is.na(oak_sod$hrs_abv25rs)] <- 0
